On particular diameter bounds for integral point sets in higher dimensions \footnote{ This work was carried out at Voronezh State University and supported by Russian Science Foundation grant 19-11-00197. }

### 2021-11-03 AvdMomZvol_rails/Article.tex
[AMZr] Edits from Usachev - p. 4

\title{
	On particular diameter bounds for integral point sets in higher dimensions
	\footnote{
		This work was carried out at Voronezh State University and supported by the Russian Science
		Foundation grant 19-11-00197.
	}
}

@@

\begin{dfn}\label{dfn1}
	A planar integral point set (PIPS) is a set $\mathcal{P}$
	of non-collinear points in plane $\mathbb{R}^{2}$ such that
	for any pair of points $P_{1}, P_{2} \in \mathcal{P}$
	the Euclidean distance $|P_{1}P_{2}|$
	between points $P_{1}$ and $P_{2}$ is integral.
\end{dfn}

@@

How can we characterize an integral point set?
First of all, we can look at its dimension;
then, we can count the number of points in it, which is always finite~\cite{anning1945integral,erdos1945integral}
and is further called the cardinality;
finally, we can naturally% TODO: naturally/essentially? cf. arxiv/1
define the diameter of a finite point set.

\begin{dfn}
	A diameter of the integral point set $\mathcal{P}$ is defined as
	\begin{equation}
		\operatorname{diam(\mathcal{P})} = \underset{P_{1}, P_{2} \in
		\mathcal{P}}{\max} |P_{1}P_{2}|
	\end{equation}
\end{dfn}

Every integral point set also has a characteristic~\cite{kemnitz1988punktmengen,kurz2005characteristic}.

\begin{dfn}
	The function $d(m, n)$ is the minimum possible diameter of
	the integral point set $\mathcal{P}$ of $n$ points in
	$m$-dimensional Euclidean space $\mathbb{R}^{m}$.
\end{dfn}

@@

For the list of known bounds for $d(m, n)$,
we refer the reader to~\cite[Theorem 1]{kurz2008bounds} or to~\cite{our-vmmsh-2018}.
We will discuss the following estimations presented at~\cite{kurz2008bounds}:
\begin{align}
	d(m, 2m + 1) \leq 8\\
	d(m, 2m + 2) \leq 13 \hypertarget{d2}\\
	d(m, 3m) \leq 109
\end{align}
and the following theorem \cite[Theorem 2.1]{kurz2008bounds}.

\begin{theorem}
	Let $\mathcal{P}$ be a planar integral point set consisting of
	$n - 2$ points on line $l_1$ and two points $P_{1}$ and $P_{2}$ on a
	parallel line $l_2$ with distance $r$ between $l_{1}$ and $l_{2}$. If there
	exist positive integers $v$, $w$ with $f^{2} + v^{2}
	= w^{2}$ and $v < 2r$, where $|P_{1}P_{2}| = f$,
	then

@@

Firstly, we discuss the classification of planar integral points sets;
then, we present some bounds for $d(m,n)$ based on planar integral point sets of particular types
and provide some general constructions for such bounds.

@@

\begin{dfn}
	A planar integral point sets of $n$ points with $n-1$ points on a straight line is called
	a \textit{facher} set.
\end{dfn}
Facher sets are very dominating examples of planar integral pont sets.
In~\cite{antonov2008maximal}, facher sets of characteristic 1 are called \textit{semi-crabs}.
For $9 \leq n \leq 122$, the diameter $d(2,n)$ is reached on a facher point set~\cite{kurz2008minimum}.

@@

\begin{dfn}
	A planar integral point sets situated in two parallel straight lines
	is called a \textit{rails} set.
\end{dfn}

@@

\begin{dfn}
	A planar integral point sets situated in two perpendicular straight lines
	is called a \textit{cross} set.
\end{dfn}
Every cross set has characteristic 1;
in~\cite{antonov2008maximal}, cross sets with only 2 points out of one of the lines are called \textit{crabs}.

\begin{dfn}
	A planar integral point sets situated in two straight lines
	that are not parallel nor perpendicular,
	is called a \textit{sciccors} set.
\end{dfn}

There is an important subclassof scissors sets.

@@

\begin{dfn}
	A planar integral point sets that is situated on a circle is called a \textit{circular}
	point set.
\end{dfn}

@@

\begin{dfn}
	A planar integral point sets that is situated on the conjunction of a circle with its center,
	is called a \textit{centered-circular} point set.
\end{dfn}

@@

\begin{theorem}
	Let $\mathcal{P}$ be a planar integral point set consisting of
	$n - k$ points on line $l_1$ and $k$ points $P_1$, $P_2$, ..., $P_k$ on a
	parallel line $l_2$ with distance $r$ between $l_1$ and $l_2$. If there
	exist positive integers $v$, $w_{ij}$, with $f_{ij}^{2} + v^{2}
	= w_{ij}^{2}$ and $v < 2r$, where $i = \overline{1, k - 1}$, $j =
	\overline{i + 1, k}$, $|P_{i}P_{j}| = f_{ij}$,
	then

@@

Below we give the planar integral point sets and the corresponding
estimates of the function $d(m, n)$ for $n = 2m + k$, $3 \leq k \leq 36$.
For convenience, we use the notation \cite{our-ped-2018,our-pmm-2018,our-vmmsh-2018}:
$\sqrt{p}/q * \{ (x_1,y_1), ...,$ $ (x_n, y_n)  \}$,
which means that each abscissa is multiplied by $1/q$
and each ordinate is multiplied by $\sqrt{p}/q$,  i.e.
$$
	\sqrt{p}/q * \{ (x_1,y_1), ..., (x_n, y_n)  \}
	=
	\left\{ \left(\frac{x_1}{q},\frac{y_1\sqrt{p}}{q}\right), ..., \left(\frac{x_n}{q},   \frac{y_n\sqrt{p}}{q}\right)  \right\}.
$$

@@

$f = 19088160$, $v = 8738$, $w = 19088162$, $\operatorname{diam(\mathcal{P})}
= 107526848$,

@@

\begin{remark}
For all the examples above, maximum in the right part of (\ref{formula1})
evaluates to the diameter of the system.
The following example shows that the maximum can be evaluated to $w$:
\end{remark}

@@

We can obtain the estimate $d(m, 2m + 2) \leq 190$ (which
is worse that known one (\hyperlink{d2}{2})).

@@

Guy gives \cite[D 20]{guy2013unsolved} a system of $8$ points located
on two parallel lines. The points of the system form a rectangle
(Figure~\ref{rectangle.pdf}):

\begin{figure}[htbp]
	\includegraphics[width=.6\linewidth]{rectangle.pdf}
	\hfill
	\includegraphics[width=.5\linewidth]{right_angle.pdf}
	\\
	\parbox{.65\linewidth}{\caption{Distances}
	\label{rectangle.pdf}}
	\hfill
	\parbox{.5\linewidth}{\caption{Counterexample}
	\label{right_angle.pdf}}
\end{figure}

@@

\section{Final remarks}
All the given PIPSs were obtained through a combination of computer search an intuition of the authors;
so, the further search may lead to better bounds employing the same consructions.

There is still no general construction for a rails or scissors PIPS of given cardinality.
For rails PIPSs, we can conjecture that there exists a set of arbitrary cardinality with 2 points on one line
and all the rest on the other;
on the other hand, we still have not found any rails PIPSs with 4 and 5 points on the lines.

>>>>

\title{
	On particular diameter bounds for integral point sets in higher dimensions
	\footnote{
		This work was carried out at Voronezh State University and supported by Russian Science
		Foundation grant 19-11-00197.
	}
}

@@

\begin{dfn}\label{dfn1}
	A planar integral point set (PIPS) is a set $\mathcal{P}$
	of non-collinear points in the plane $\mathbb{R}^{2}$ such that
	for any pair of points $P_{1}, P_{2} \in \mathcal{P}$
	the Euclidean distance $|P_{1}P_{2}|$
	between points $P_{1}$ and $P_{2}$ is integral.
\end{dfn}

@@

How can we characterize an integral point set?
First of all, we can look at its dimension;
then, we can find its cardinality, which is always finite~\cite{anning1945integral,erdos1945integral};
finally, we can naturally % TODO: naturally/essentially? cf. arxiv/1
define the diameter of a finite point set.

\begin{dfn}
	The diameter of an integral point set $\mathcal{P}$ is defined as
	\begin{equation}
		\operatorname{diam(\mathcal{P})} = \underset{P_{1}, P_{2} \in
		\mathcal{P}}{\max} |P_{1}P_{2}|
		.
	\end{equation}
\end{dfn}

Every integral point set has also a characteristic~\cite{kemnitz1988punktmengen,kurz2005characteristic}.

\begin{dfn}
	The function $d(m, n)$ is the minimal possible diameter of
	the integral point set $\mathcal{P}$ of $n$ points in
	$m$-dimensional Euclidean space $\mathbb{R}^{m}$.
\end{dfn}

@@

For the list of known bounds for $d(m, n)$,
we refer the reader to~\cite[Theorem 1]{kurz2008bounds} or to~\cite{our-vmmsh-2018}.
We will discuss the following ones presented at~\cite{kurz2008bounds}:
\begin{align}
	d(m, 2m + 1) \leq 8\\
	d(m, 2m + 2) \leq 13 \label{eq:d_m_2m+2}\\
	d(m, 3m) \leq 109
\end{align}
and the following theorem \cite[Theorem 2.1]{kurz2008bounds}.

\begin{theorem}
	Let $\mathcal{P}$ be a planar integral point set consisting of
	$n - 2$ points on the line $l_1$ and two points $P_{1}$ and $P_{2}$ on a
	parallel line $l_2$ with distance $r$ between $l_{1}$ and $l_{2}$. If there
	exist positive integers $v$, $w$ with $f^{2} + v^{2}
	= w^{2}$ and $v < 2r$, where $|P_{1}P_{2}| = f$,
	then

@@

First, we discuss the classification of planar integral points sets;
then, we present some bounds for $d(m,n)$ based on planar integral point sets of special types
and provide some general constructions of such bounds.

@@

\begin{dfn}
	A planar integral point sets of $n$ points with $n-1$ points on a straight line is called
	a \textit{facher} set.
\end{dfn}
Facher sets are very dominating examples of planar integral point sets.
In~\cite{antonov2008maximal}, facher sets with characteristic 1 are called \textit{semi-crabs}.
For $9 \leq n \leq 122$, the diameter $d(2,n)$ is reached on a facher point set~\cite{kurz2008minimum}.

@@

\begin{dfn}
	A planar integral point set situated in two parallel straight lines
	is called a \textit{rails} set.
\end{dfn}

@@

\begin{dfn}
	A planar integral point set situated in two perpendicular straight lines
	is called a \textit{cross} set.
\end{dfn}
Every cross set has characteristic 1;
in~\cite{antonov2008maximal}, cross sets with only 2 points out of one of the lines are called \textit{crabs}.

\begin{dfn}
	A planar integral point set situated in two straight lines
	that are not parallel nor perpendicular,
	is called a \textit{scissors} set.
\end{dfn}

There is an important subclass of scissors sets.

@@

\begin{dfn}
	A planar integral point set that is situated on a circle is called a \textit{circular}
	point set.
\end{dfn}

@@

\begin{dfn}
	A planar integral point set that is situated on the conjunction of a circle with its center,
	is called a \textit{centered-circular} point set.
\end{dfn}

@@

\begin{theorem}
	Let $\mathcal{P}$ be a planar integral point set consisting of
	$n - k$ points on the line $l_1$ and $k$ points $P_1$, $P_2$, ..., $P_k$ on a
	parallel line $l_2$ with distance $r$ between $l_1$ and $l_2$. If there
	exist positive integers $v$, $w_{ij}$, with $f_{ij}^{2} + v^{2}
	= w_{ij}^{2}$ and $v < 2r$, where $i = \overline{1, k - 1}$, $j =
	\overline{i + 1, k}$, $|P_{i}P_{j}| = f_{ij}$,
	then

@@

Below we give the planar integral point sets and the corresponding
estimates of the function $d(m, n)$ for $n = 2m + k$, $3 \leq k \leq 36$.
For the sake of brevity, we use the notation \cite{our-ped-2018,our-pmm-2018,our-vmmsh-2018}:
$\sqrt{p}/q * \{ (x_1,y_1), ...,$ $ (x_n, y_n)  \}$,
which means that each abscissa is multiplied by $1/q$
and each ordinate is multiplied by $\sqrt{p}/q$,  that is
$$
	\sqrt{p}/q * \{ (x_1,y_1), ..., (x_n, y_n)  \}
	=
	\left\{ \left(\frac{x_1}{q},\frac{y_1\sqrt{p}}{q}\right), ..., \left(\frac{x_n}{q},   \frac{y_n\sqrt{p}}{q}\right)  \right\}.
$$

@@

$f = 19088160$, $v = 8738$, $w = 19088162$, $\operatorname{diam}(\mathcal{P})
= 107526848$,

@@

\begin{remark}
In all the examples above, the maximum in the right-hand side of (\ref{formula1})
is attained on the diameter of the system.
The following example shows that the maximum can be attained on $w$:
\end{remark}

@@

We can obtain the estimate $d(m, 2m + 2) \leq 190$ (which
is weaker than the one given in~\eqref{eq:d_m_2m+2}.

@@

Guy gives \cite[D 20]{guy2013unsolved} a PIPS of $8$ points located
on two parallel lines. The points of the set form a rectangle
(Figure~\ref{rectangle.pdf}):

\begin{figure}[htbp]
	\includegraphics[width=.6\linewidth]{rectangle.pdf}
	\hfill
	\includegraphics[width=.5\linewidth]{right_angle.pdf}
	\\
	\parbox{.65\linewidth}{\caption{Distances}
	\label{rectangle.pdf}}
	\hfill
	\parbox{.5\linewidth}{\caption{Contradiction}
	\label{right_angle.pdf}}
\end{figure}

@@

\section{Final remarks}
All the discussed PIPSs were obtained through a combination of computer search and intuition of the authors;
so, the further search may lead to better bounds employing the same constructions.

There is still no general construction for a rails or scissors PIPS of arbitrary cardinality.
For rails PIPSs, we can conjecture that there exists a set of arbitrary cardinality, with 2 points on one line
and all the rest on the other;
on the other hand, we still have not found any rails PIPSs with 4 and 5 points on the lines.
